---
title: "Tour du Capitol Reef"
subtitle: 'guide de cours: *to be updated with more details*'
author: "Buchanan Kerswell"
date: "9/13/2020"
output:
  rmdformats::readthedown:
    css: style.css
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(knitr)
library(kableExtra)
library(grid)
library(gridExtra)
library(patchwork)
library(plotKML)
library(ggmap)
library(cowplot)
library(geosphere)
library(rmdformats)
```

```{r plots, include=FALSE, message=FALSE}
e <- readGPX('./route.gpx')
e <- e$tracks[[1]]$`Tour du Capitol Reef` %>%
  tibble() %>%
  mutate('ele' = as.numeric(ele),
         dist = distHaversine(cbind(lon, lat),
                                cbind(lag(lon), lag(lat)))) %>%
  mutate(
    'dist' = dist %>% replace_na(0),
    cumdist = cumsum(dist)
  )
lat <- c(min(e$lat), max(e$lat))
lon <- c(min(e$lon), max(e$lon))
bbox <- make_bbox(lon, lat)
b <- get_map(bbox, maptype = 'terrain', source = 'google')
dut <- map_data('state', region = 'utah')
bboxut <- make_bbox(c(min(dut$long), max(dut$long)), c(min(dut$lat), max(dut$lat)))
but <- get_map(bboxut, maptype = 'terrain', source = 'google', color = 'bw')
p1inset <- ggmap(but) +
  geom_path(data = dut, aes(x = long, y = lat), size = 0.5) +
  geom_rect(data = tibble(lat = lat,
                          lon = lon),
            aes(xmin = lon[1], xmax = lon[2], ymin = lat[2], ymax = lat[1]),
            fill = 'deeppink4',
            alpha = 0.1,
            color = 'black',
            size = 0.25) +
  ylab('') +
  xlab('') +
  theme_map()
p1 <- ggmap(b) +
  geom_point(data = e, aes(lon, lat, color = ele*3.28)) +
  scale_color_gradient(name = 'Elevation\n(ft)', low = 'chocolate4', high = 'lemonchiffon') +
  ylab('') +
  xlab('') +
  theme_bw(base_size = 11) +
  theme(
    panel.border = element_rect(size = 2),
    axis.ticks = element_line(size = 1)
  ) +
  geom_label(aes(x = -111.25, y = 38.5, label = 'Tour du\nCapitol Reef\n231 mi\n17.35k (ft)')) +
  inset(ggplotGrob(p1inset), xmin = -110.95, xmax = -110.55, ymin = 37.75, ymax = 38.18)
p1
ggsave('./figs/route.png', device = 'png', plot = p1)

p2 <- e %>%
  ggplot() +
  geom_area(aes(x = cumdist*0.000621371, y = ele*3.28), alpha = 0.5) +
  xlab('Distance (mi)') +
  ylab('Elevation (ft)') +
  scale_x_continuous(expand = c(0,0)) +
  theme_minimal(base_size = 11)
p2
ggsave('./figs/ele.png', device = 'png', plot = p2)
```

## The route {.flexbox .vcenter}

```{r map, echo=FALSE, out.width='100%', fig.cap='Figure 1. Map of Tour du Capitol Reef. The full segment can be found on [komoot](https://www.komoot.com/tour/257901775). Data for map from `google`'}
p1
```

## The profile {.flexbox .vcenter}

```{r profile, echo=FALSE, out.width='100%', fig.cap='Figure 2. Elevation profile of Tour du Capitol Reef. The full segment can be found on [komoot](https://www.komoot.com/tour/257901775)'}
p2
```

## The experience {.flexbox .vcenter}

The route passes through pristine UT desert, including flybys of:

1. [Mars Desert Research Station](http://mdrs.marssociety.org/)
2. [Goblin Valley State Park](https://stateparks.utah.gov/parks/goblin-valley/) & the [San Rafael Swell](https://utah.com/san-rafael)
3. Traversing much of the [Waterpocket fold](https://capitolreef.org/blog/capitol-reef-the-waterpocket-fold/)
4. [Scenic Byway 12](https://www.visitutah.com/articles/the-all-american-road-scenic-byway-12/) including the infamous **hogsback segment**, or perhaps to [Hells Backbone](https://utah.com/grand-staircase-escalante/hells-backbone)
5. Beautiful Torrey and [Capitol Reef National Park](https://www.nps.gov/care/index.htm)
6. Finale on a single track to [Cassidy Arch](https://utah.com/hiking/capitol-reef-national-park/cassidy-arch-trail) and the [Grand Wash](https://www.tripadvisor.com/Attraction_Review-g143017-d212116-Reviews-Grand_Wash_Trail-Capitol_Reef_National_Park_Utah.html)

#### ... and many more

## The details {.flexbox .vcenter .tabset}

### Summary
```{r table, echo=FALSE, message=FALSE}
d <- tibble(
  day = factor(c(1, 2, 3, 4)),
  depart = c('SR24', 'Sunlit Oasis', 'Boulder UT', 'Torrey UT'),
  arrive = c('Sunlit Oasis', 'Boulder UT', 'Torrey UT', 'Veyo UT'),
  `distance (mi)` = c(60.1, 65.9, 54.7, 50.3),
  `climbing (ft)` = c(1946, 5256, 4751, 5397),
  `descent (ft)` = c(2392, 3832, 4741, 6785)
) %>%
  mutate(
    `distance (km)` = `distance (mi)` * 1.609,
    `climbing (m)` = `climbing (ft)` / 3.28,
    `descent (m)` = `descent (ft)` / 3.28,
    `ETA (hr)` = `distance (mi)` / 10
  )

d %>%
  select(c(1:6, 10)) %>%
  kbl(digits = 0) %>%
  kable_styling(
    bootstrap_options = c('hover', 'striped'),
    font_size = 18,
    position = 'center',
    full_width = TRUE
  )
```

### Detailed agenda

#### A note about parking

This course is a lolipop, which begins north of Caineville, UT, and also finishes in Caineville. We should be able to leave a vehicle in Caineville, which we will return to on November 25 after three days of riding. *This means that we need to figure out a way to get to the start line, at the junction of HW24 & Temple Mt Road.* We can sort this detail out later...

#### Day 0 Saturday Nov 21

##### Afternoon (1700?)

- Buchanan arrives in SLC
- Meet up
- Gear check

##### Evening

- Final gear check
- Prepare for Grand Depart

#### Day 1 Sunday Nov 22

##### 0700

##### 0715 Sunrise

##### 0730

- Depart for the Cainevilee, Utah (3.5 hr drive)

<center>

<iframe src="https://www.google.com/maps/embed?pb=!1m14!1m8!1m3!1d12466.943041453085!2d-110.5840104!3d38.6319595!3m2!1i1024!2i768!4f13.1!3m3!1m2!1s0x0%3A0x0!2zMzjCsDM3JzUwLjgiTiAxMTDCsDM0JzAwLjIiVw!5e0!3m2!1sen!2sus!4v1600108354228!5m2!1sen!2sus" width="400" height="300" frameborder="0" style="border:0;" allowfullscreen="" aria-hidden="false" tabindex="0"></iframe>

</center>

##### 1100

- Park vehicle in Cainevlille, Utah and travel to the junction of HW24 & Temple Mt Road
- Kit up and ride approx. 6 hr to Sunlit Oasis

<center>

<iframe src="https://www.google.com/maps/embed?pb=!1m14!1m8!1m3!1d12537.899959683604!2d-111.119681!3d38.222084!3m2!1i1024!2i768!4f13.1!3m3!1m2!1s0x0%3A0x59f1481fffb8e2f4!2sSunlit%20Oasis!5e0!3m2!1sen!2sus!4v1600108283572!5m2!1sen!2sus" width="400" height="300" frameborder="0" style="border:0;" allowfullscreen="" aria-hidden="false" tabindex="0"></iframe>

</center>

##### 1900

- Arrive at Sunlit Oasis
- Check in

##### 1915 Sunset

- Dinner (@ Sunlit Oasis?)
- Debauch

#### Day 2 Monday Nov 23

##### Sunrise 0715

- Breakfast at Sunlit Oasis (included)
- Kit up and ride approx. 6.5 hr to Boulder, UT

<center>

<iframe src="https://www.google.com/maps/embed?pb=!1m18!1m12!1m3!1d37261.95641931418!2d-111.45143697028186!3d37.89707148787745!2m3!1f0!2f0!3f0!3m2!1i1024!2i768!4f13.1!3m3!1m2!1s0x0%3A0x0!2zMzfCsDU0JzExLjciTiAxMTHCsDI1JzE5LjAiVw!5e0!3m2!1sen!2sus!4v1600108790769!5m2!1sen!2sus" width="400" height="300" frameborder="0" style="border:0;" allowfullscreen="" aria-hidden="false" tabindex="0"></iframe>

</center>

##### 0830

- Depart for Boulder, UT

##### 1500

- Arrive in Boulder, UT
- Check into hotel
  * [Boulder Mt Lodge](https://www.google.com/maps/place/Boulder+Mountain+Lodge/@37.9006605,-111.4288698,14.77z/data=!4m12!1m3!2m2!1shotel!6e3!3m7!1s0x8735f47c57c09f47:0x79f0dc7ff24be145!5m2!4m1!1i2!8m2!3d37.9020588!4d-111.4245629)
  * [Circle Cliffs Motel](https://www.google.com/maps/place/Circle+Cliffs+Motel/@37.9006605,-111.4288698,14.77z/data=!4m12!1m3!2m2!1shotel!6e3!3m7!1s0x0:0xe2a94c7ff0cd6f27!5m2!4m1!1i2!8m2!3d37.9066894!4d-111.4245361)
- Dinner in town
  * [Hell's Backbone Grill](https://www.google.com/maps/place/Hell's+Backbone+Grill+%26+Farm/@37.9020607,-111.4265417,17z/data=!3m1!4b1!4m5!3m4!1s0x8735f4581f7cdeb3:0x8c2850804cb69ac6!8m2!3d37.9020569!4d-111.4245623)
  * [Burr Trail Grill](https://www.google.com/maps/place/Burr+Trail+Grill/@37.9020607,-111.4265417,17z/data=!4m5!3m4!1s0x0:0xcf09f9761cb9f2d1!8m2!3d37.9014265!4d-111.4246045)
  * *Or cook ourselves!*
  
##### 1915 Sunset

- Debauch

#### Day 3 Tuesday Nov 24

##### 0715 Sunrise

##### 1915 Sunset


#### Day 4 Wednesday Nov 25

##### 0715 Sunrise

- Ride in the park for approx. 5 hr
- Ride to Caineville to collect vehicle

##### 1915 Sunset

- Drive to Veyo, UT for Cabin festivities

## Route summary {.flexbox .vcenter}

```{r summary, echo=FALSE, message=FALSE, out.width='100%', fig.cap='Figure 3. Important route information summarized by day. '}
s <- d %>%
  select(where(is.numeric)) %>%
  pivot_longer(cols = everything(), names_to = 'vars', values_to = 'vals') %>%
  group_by(vars) %>%
  summarise(total = round(sum(vals), 1),
            avg = round(mean(vals)), 1)
p <- d %>%
  select(day, `distance (mi)`, `climbing (ft)`, `descent (ft)`, `ETA (hr)`) %>%
  pivot_longer(-day, names_to = 'vars', values_to = 'vals') %>%
  group_by(vars) %>%
  group_by(day) %>%
  ggplot() +
  geom_area(aes(x = day, y = vals, group = vars, fill = vars), color = 'black', alpha = 0.5, show.legend = FALSE) +
  geom_text(data = s %>% filter(vars %in% c('ETA (hr)', 'climbing (ft)', 'descent (ft)', 'distance (mi)')), aes(x = 2.5, y = 0, group = vars, label = paste0('avg: ', avg, '/day', '\ntotal: ', total)), vjust = -1, size = 3) +
  ylab('') +
  scale_x_discrete(expand = c(0,0)) +
  scale_fill_brewer(palette = 'Paired') +
  facet_wrap(~ vars, scales = 'free_y') +
  theme_bw(base_size = 11)
p
```

## The map {.flexbox .vcenter}

<center>

<iframe src="https://www.google.com/maps/d/u/0/embed?mid=1jU5P6ApFli7oH5DRchvhPMoNPf_VNUtl" width="640" height="480"></iframe>

</center>

## The gallery {.flexbox .vcenter}

```{r gallery, echo=FALSE, message=FALSE, fig.align='center'}
fnames <- list.files('./figs')
for(i in fnames){
  p <- ggdraw() +
    draw_image(paste0('./figs/', i))
  print(p)
}
```